---
title: "Week 6 HW - Who were the winners?"
author: "Jordan Sibley" 
date: "May 15, 2025"
format: html
editor_options: 
  chunk_output_type: inline
---

## Directions 

At the conclusion of the ASDN project the PIs decided to hand out first, second, and third prizes to the observers who measured the most eggs. Who won? Please use R and dbplyr to answer this question, and please submit your R code. Your code should print out:


```
# Ordered by: desc(total_eggs)
  Name            total_eggs
  <chr>                <int>
1 Vanessa Loverti        163
2 Dylan Kessler           87
3 Richard Lanctot         50
```


You’ll want to load database tables using statements such as:

`egg_table <- tbl(conn, "Bird_eggs")`

and then use tidyverse grouping, summarization, joining, and other functions to compute the desired result.

Also, take your final expression and pipe it into `show_query()`. If you used multiple R statements, did dbplyr create a temporary table, or did it manage to do everything in one query? Did it limit to the first three rows using an R expression or an SQL LIMIT clause?


```{r}
# Load libraries
library(dbplyr)
library(tidyverse)
library(DBI)
library(duckdb)
```


```{r}
# Establish databse connection 
conn <- DBI::dbConnect(duckdb::duckdb(), dbdir="database/database.db")
```


```{r}
# View tables 
DBI::dbListTables(conn)
```

```{r}
# Read in and view tables
personnel <- tbl(conn, "Personnel")
personnel

bird_eggs <- tbl(conn, "Bird_eggs")
bird_eggs

camp_assignment <- tbl(conn, "Camp_assignment")
camp_assignment

site <- tbl(conn, "Site")
site

bird_nests <- tbl(conn, "Bird_nests")
bird_nests
```

```{r}
# Query to determine who measured the most eggs

winners <- bird_eggs %>%
  left_join(bird_nests, by = c("Year" = "Year", "Site" = "Site", "Nest_ID" = "Nest_ID")) %>%
  left_join(personnel, by = c("Observer" = "Abbreviation")) %>% 
  group_by(Name) %>% 
  summarize(total_eggs = n()) %>% 
  arrange(desc(total_eggs)) %>% 
  head(3)

winners %>% 
  show_query()

winner_results <- winners %>% 
  collect()

print(winner_results)
```

I was able to answer the question in one dbplyr query, The observer who counted the most eggs was Vanessa Loverti with 163 
